---
title: "Advanced Statistical Inference HW1"
author: 'Dijia Tang'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Bronze Tier

Each task is worth 2 points.

## Task 1

Use the attrition data from the rsample package.

```{r}
library(rsample)
library(Hmisc)
data("attrition")
summary(attrition)
describe(attrition)

```

Engage in some data exploration by using `summary` functions.

In addition to the `summary` function, the `describe` functions from the psych and Hmisc packages are great (but for slightly different purposes). Use any function you wish

## Task 2

How do those variables look? Do they appear to all make sense? Are there any extreme values? Are any of them coded in such a way that don't make sense? Discuss anything interesting that you found.

Be sure to use the `str` function or the `glimpse` function from dplyr to check the structure of the data.

1. The StockOptionLevel variable is hard to interpret without any given contexts. 
2. The maximum number in DistanceFromHome is quite extreme. Normally, people wouldn't travel 29 miles everyday to go to work. However, as we can't be sure if mile is used as the unit. It may be too early to draw this conclusion. 
3. BusinessTravel should be coded as ord.factor since it represents the travel frequency.
4. The maximum value in YearsSinceLastPromotion can be extreme. It doesn't make much sense if one stays in a company for 15 years without any promotion. Same applies to YearsInCurrentRole and YearsWithCurrManager.
5. The mean of MonthlyIncome is 6503 while that of MonthlyRate is 14313. There is a vast difference between them. Although we don't know the exact difference between these two variables. Intuitively, there shouldn't be such a huge difference.
6. The max DailyRate is 1499 while the max HourlyRate is 100. How can someone make 1499 a day when his hourly rate is only 100? This cannot make sense.
```{r}
library(tidyverse)
str(attrition)
glimpse(attrition)
table(attrition$PerformanceRating)
```

## Task 3

Pick at least 3 variables for which you want to explore relationships. Given any issues you found in Task 2, convert variables as needed with `as.factor` (forces a variable into a factor -- think about a nominal variable), `as.numeric` (forces a variable into a numeric), or anything else that you deem necessary (e.g., create new variables with `mutate`). 

attrition = attrition %>% 
mutate(BusinessTravel = factor(BusinessTravel, levels=c("Non-Travel", "Travel_Rarely", "Travel_Frequently"), ordered=TRUE))

***Variable Picked***: Age, YearsAtCompany, MonthlyRate, PercentSalaryHike, YearsSinceLastPromotion

## Task 4

Using the ggplot2 package, create some visualizations. The structure of a ggplot object is as follows:

```{r}
attrition %>% 
ggplot(aes(x = Age, y = YearsAtCompany)) + 
  geom_point()
```

If you want to add some type of grouped function (by color, for instance), you might do something like this:

```{r}
attrition %>% 
ggplot(., aes(x = Age, y = YearsAtCompany, color = Gender)) + 
  geom_point()
```

Once you have produced a visualization, explain what it means. Naturally, good viz doesn't need explanation, but let's do it regardless!

***Explanation***: We can see that the in age range 20 to 40, people tend to stay at the company for a pretty long time, especially male employees. However, after age 40, employees tend to leave the company, as the data points become more sparse.

## Task 5

Use the `simViz` function from inferviz to see how well you actually do in finding an actual relationship.

The first graphy below is the actual relationship.
```{r}
if(!(require(devtools))) {install.packages("devtools")}
devtools::install_github("saberry/inferviz")
library(inferviz)

simViz(attrition, Age, YearsAtCompany, distractors = 3, answer = FALSE)

```


First, pick two new variables for relationship exploration -- make sure that you have not seen them before.

When you run it the first time, keep the distractors argument at the default value. When asked if you want to see the real data select 1 (No). Now try it with more distractors (maybe 5 or 6).

How did you do? Where you able to spot the real relationship or were you fooled by your basic human need for pattern finding?

Discuss what you think this means.

***Discussion***: There is not a very obvious pattern to be distinguished. All graphs seem to show that there is a better chance to get a salary hike shorter after a promotion than longer. It is hard to identify the actual relationship.
```{r}
simViz(attrition, YearsSinceLastPromotion, PercentSalaryHike, distractors = 3, answer = F)
simViz(attrition, YearsSinceLastPromotion, PercentSalaryHike, distractors = 5, answer = F)
simViz(attrition, YearsSinceLastPromotion, PercentSalaryHike, distractors = 6, answer = F)
```

# Silver Tier

For up to 4 points, use the leadershipRatingsAgreement.csv file in the data folder. 

You can use any of the following functions to read a csv file into R (from slow to fast):

```{r}
lead <- data.table::fread("C:/Users/dijia/OneDrive/ND/ITAO 70200 Adv Stats/leadershipRatingsAgreement.csv")
glimpse(lead)
library(DT); library(broom)
leadcor = cor(lead[,2:8], method = "pearson", use = "complete.obs")
knitr::kable(round(leadcor, 2))

```


Which variables have the strongest positive relationship and which have the strongest negative relationship? Once you have found those relationships, plot them both.

***Answer***: Productivity and Vitality have the strongest positive relationship.
Vitality and Effect have the strongest negative relationship.
```{r}
lead %>% 
ggplot(aes(x = productivity, y = vitality))+
  geom_point() +
  geom_smooth(method = 'lm',se=F)+
  theme_light()

lead %>% 
ggplot(aes(x = vitality, y = effect))+
  geom_point() +
  geom_smooth(method = 'lm',se = F, color = 'red')+
  theme_light()
```

Which 4 variables have the strongest relationship with the effect variable? Explain how you found this relationship. Produce plots for each of those relationships and include both linear trend lines and smoothed trend lines.  

***Explanation***: The variables are Enabling, Forceful, Strategic, and Operational. I found these relationships by looking at the effect column of the correlation matrix created earlier and comparing the absolute values of the correlation values. 
```{r}
lead %>% 
ggplot(aes(x = enabling, y = effect))+
  geom_point() +
  geom_smooth(method = 'lm',se = F, color = 'red')+
  geom_smooth(se = F)+
  theme_light()

lead %>% 
ggplot(aes(x = forceful, y = effect))+
  geom_point() +
  geom_smooth(method = 'lm',se = F, color = 'red')+
  geom_smooth(se = F)+
  theme_light()

lead %>% 
ggplot(aes(x = strategic, y = effect))+
  geom_point() +
  geom_smooth(method = 'lm',se = F, color = 'red')+
  geom_smooth(se = F)+
  theme_light()

lead %>% 
ggplot(aes(x = operational, y = effect))+
  geom_point() +
  geom_smooth(method = 'lm',se = F, color = 'red')+
  geom_smooth(se = F)+
  theme_light()

 
```

# Gold Tier

For up to 6 points, use the leadership data found in the data folder. This is a .dta file (from Stata) and can be read in using the appropriate function from the haven package. 

```{r}
ship <- haven::read_dta("C:/Users/dijia/OneDrive/ND/ITAO 70200 Adv Stats/leadership")
glimpse(ship)
```

Using verbs from the dplyr package (specifically, `group_by` and `summarize`), find the *effect* mean for each leaderID. Which leader(s) had the highest effect mean and which leader(s) had the lowest effect mean.
```{r}
ef <- ship %>% 
  group_by(leaderID) %>% 
  summarise(effectmean = mean(effect)) 
```
Sort from Lowest
**Leader 1589 has the lowest effect mean.**
```{r}

ef %>% 
  arrange(effectmean)

```
Sort from Highest
**Leader 1644, 2507, and 8467 all have the highest effect mean, which is 10.**
```{r}
ef %>% 
  arrange(desc(effectmean))
```


Take your highest and lowest leaders and keep the leaderID. With the leaderID in hand, find those leaders in the leadershipRatingsAgreement data.
```{r}
selected_leaders <- data.frame(leaderID = c(1644,2507,8467,1589))
find <- merge(selected_leaders, lead, by.x = 'leaderID', by.y = 'leaderID')
find <- find %>% 
  select(-effect) %>% 
  pivot_longer(-leaderID, names_to = "style", values_to = "value")

find %>% 
  ggplot()+
  geom_point(aes(x = style, y=value, color = factor(leaderID))) 
  
```

Excluding effect, construct a plot with forceful, enabling, strategic, operational, vitality, and productivity on the x-axis (the y-axis will contain the value for each variable). Next color the points by leaderID.