---
title: "Mixed_models_raneff"
author: "Dijia Tang"
date: "12/3/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages

```{r}
library(tidyverse)
library(lme4)
library(merTools)
```

Load data and linear model

```{r}
teamdata <- miceadds::load.Rdata2('teamPerc.RData')
team <- teamdata
glimpse(team)
linModteam <- lm(effect~leader_age+takesCharge+declares+pushes,data=team)
summary(linModteam)
```

Randome effect

```{r}
raneffect <- lmer(effect ~ 1 +  (1|leaderID), data = team)
summary(raneffect)
.2923/(.2923+1.3298)
```
We can see that the ICC is 0.18. This means that leaderID's random effect would explain 18% of the variace in the dependent variable.
```{r}
plotREsim(REsim(raneffect), labs = TRUE)
```
Plotting the random effect, we can see that the effect cab be pretty significant for some leaders.

Add the random effect back to the linear model. 

```{r}
ranlinMod <- lmerTest::lmer(effect ~ leader_age+takesCharge+declares+pushes + (1|leaderID), data = team)
summary(ranlinMod)
```

Add predictions

```{r}
mixedPred <- predict(ranlinMod)

slimPred <- predict(linModteam)

allPred <- cbind(actual = team$effect, 
                 mixed = mixedPred, 
                 slim = slimPred)
```

Draw the plot for the linear and the mixed models

```{r}
par(mfrow=c(1,2))
plot(allPred[, "actual"], allPred[, "slim"])
plot(allPred[, "actual"], allPred[, "mixed"])
```

We can see that the linear model is denser. So in this case, our linear model predicts effect better than the mixed model.

